Reactive silver inks for antiviral, repellent medical textiles with ultrasonic bleach washing durability compared to silver nanoparticles

### 2022-01-20 Latex/Response Letter1.tex

% request for removal from telemarketing lists

\documentclass[12pt]{letter}

\newcommand{\cX}{{\mathcal X}}

\usepackage{gensymb}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
%\usepackage{hyperref}
\usepackage[normalem]{ulem}
\usepackage{textcomp}
\usepackage{color}
\usepackage{xcolor}
%\usepackage{pdfcomment}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pdfcomment}

\usepackage{paralist}

\usepackage{setspace}
\usepackage{hyperref}
%\usepackage{color}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{ulem}
\usepackage{graphicx}

%\definecolor{blue}{rgb}{0,0,0.543}
%\definecolor{blue}{rgb}{0,0.08,0.45}
%\def\blue[rgb]{0 0 139}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=5.9in
\topmargin=-.5in
\textheight=9in

\usepackage{xargs}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}

%\newcommand{\blue}[1]{\textcolor{blue}{#1}} %for displaying red texts
%\newcommand{\strikeout}[1]{\textcolor{blue}{\sout{#1}}} %for displaying red texts

\newcommandx{\commentPaul}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}

\defineavatar{Paul}{color=red}

%\newcommandx{\commentPaul}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}

\newcommand{\strikeout}[1]{\textcolor{blue}{\sout{#1}}} %for displaying red texts

\newcommand{\red}[1]{\textcolor{red}{#1}} %for displaying red texts
\newcommand{\blue}[1]{\textcolor{blue}{#1}} %for displaying red texts
\newcommand{\Hmax}{\hat{H}}

\onehalfspacing
%\name{Sajad Haghanifar}
\name{Paul W Leu}

\date{\today}

\begin{document}

\begin{letter}{}

\opening{Dear Dr.~Ned Bowden:}

We thank the three reviewers for their evaluation, generally positive comments and conclusion that our work is of interest 
to the readership of ACS Applied Nano Materials.

% We thank the three reviewers for their evaluation of our manuscript and comments.  
%We thank the reviewers' statements that ``this review article is comprehensive to introduce the various sources of bio-inspiration, the critical performance requirements and the strategies that have been carried out to tackle the challenges" and we have provided ``an elaborate and comprehensive arrangement of various transparent substrates and barrier layers for optoelectronics."

We have made improvements to our paper based on the reviewers' comments and suggestions.
Below, we summarize the changes that have been made and respond to the reviewers' comments.
Considering the importance of the topic of this
article on 
%for application of 
durable, anti-viral fabric coatings,
we resubmit this manuscript for your consideration of publication in ACS Applied Nano Materials.
Thank you for your consideration.

\closing{My best regards,}

%Paul Leu

\end{letter}

\newpage

\textbf{Editor}

\begin{enumerate}

\item 
Please add a 
scheme near the 
beginning (probably 
defined as Scheme 1) 
to illustrate the 
system 
(structure/chemical 
composition) and the 
synthesis/fabrication 
from a chemistry 
perspective.

\blue{A schematic showing the synthesis and structure of the functionalized nano graphene oxide has been added to Figure 1.}

\item 
Please remove the box 
around the TOC graphic and 
increase its clarity/resolution.

\blue{The box has been removed and the resolution of the graphic has been improved by using a .tif file.}

\item In Figures 1c, 5c, S5, and 2c please use a.u. for the units and not use A.u.

\blue{This has been corrected.}

\item The contact angle data in Figures 3a and 5e are reported with too much precision (i.e., too many significant figures). Contact angles should be reported as xx degrees not xx.y degrees. Please check carefully and revise where needed.

\blue{This has been corrected.}

\item Please check Figure S4, some of it is missing.

\blue{This has been corrected.}

\item Please reformat the Supporting Information so that each figure appears on the same page as its figure caption.

\blue{This has been corrected.}

\end{enumerate}

\newpage

\textbf{Reviewer: 1}

\textbf{Recommendation:} Publish after minor revisions noted.

This manuscript reports a new strategy for bleach washable and durable anti-virofouling fabiric coating using functionalized graphene oxide derived from coals. This strategy enables to easily endow an antiviral property with the fabric used here based on simple and cost-effective processes. The coated functionalized graphene oxides attached on the surface of the fabrics show good anti-virofouling property and durability. In my view, the results in this manuscript is publishable in this prestigious journal after major revision.

\blue{We thank the reviewer for his or her positive comments and evaluation of our article to be publishable in ACS Applied Nano Materials.
We have made improvements based on the reviewer's comments as shown below.}
%\commentPaul{Think all page numbers and paragraph numbers need to be checked and changed as I think they have changed after modifying figures.}

% to introduce the various sources of bio-inspiration,the critical performance requirements and the strategies that have been carried out to tackle the challenges" and the content we provide is “solid and well drafted". We have made improvement based on reviewer's comments:}

\begin{enumerate}
\item The coal-derived graphene oxide has $\sim$8 at.\% of oxygen content. Actually, this amount of oxygen content is not sufficient to call this material graphene oxide. I am not raising a big issue on the this “graphene oxide”. But, if the authors can search a more proper name for their materials, it can be changed.

\blue{The low oxygen content is due to the functionalization of the graphene oxide (GO). The graphene oxide oxygen content before functionalization is about 34.5\%, but after functionalization much of the oxygen content is lost due to the reaction to form amide between carboxylic groups on the GO and the amines.
%and the loss of 
% there is loss of O (formation of water between COOH and NH$_2$), etc., and so you’ll see a decreased oxygen content in the functionalized GO. 
This explanation has been added to the end of page 7, line 15-20. 
%\commentPaul{Add verbatim the text that has been added so the reviewers don't have to go searching for it.}
Additionally, the text has been corrected throughout the paper to refer to the nanomaterial as functionalized nano-graphene oxide for clarity.}
%\commentPaul{In some places, you use nano graphene oxide.  In others nano-graphene oxide.  It needs to bee consistent.  Please check that ``nano-graphene oxide" is used throughout.}
%\commentPaul{Perhaps, we can just call it graphene oxide?  I'm not sure why we want the prefix nano-?}

\item The mechanism for antiviral property of the GO-ODA need to be explained. Does the anti-fouling or antiviral properties of the materials originate from oxygen function group or ODA functional group?

\blue{ The properties come from the ODA functional group and nGO nanostructure. The discussion on page 20, paragraphs 1-3 have been modified for clarity. 
%\commentPaul{Add verbatim the last to paragraphs so the reviewers don't have to go searching for it.}
The anti-viral behavior observed 
%is anti-virofouling which 
comes from the repellency and inactivation property of the ODA functional group combined with the nanoroughness from the nGO.}
%The virus may not have been killed, but you we can accurately state that infectivity of the virus was reduced. This is just as good as killing the virus. 
%When we dilute the virus out from the original samples, there is no increase in virus in higher dilutions compared to decreased virus in the lower dilutions suggesting a virus inhibition of infectivity effect. }

\item How about the anti-fouling or antiviral property of GO-ODA compared with conventional GOs? Please explain the advantages of GO-ODA compared to conventional GO or nitrogen doped GO.

\blue{The following text was added to page 4, paragraph 2 ``nGO-ODA offers the advantage of simultaneous liquid repellency and viral inhibition properties compared to conventional GO or nitrogen doped GO, which only demonstrate virus inhibition."}

\end{enumerate}

\newpage
\textbf{Reviewer: 2}

\textbf{Recommendation:} Major revisions needed as noted.

\blue{We thank the reviewer for his or her constructive comments. % and evaluation of our article to be published in this prestigious journal.
We have made improvements based on the reviewer's comments as shown below.}

\begin{enumerate}

\item    Page 9 line 8 “This provides ………. wetting property.” It seems like roughness is provided for the dip coating of nGO-ODA inspite of superhydrophobic wetting property. Please confirm as optical image in fig 3c ii contradicts this statement.

\blue{The paragraph on page 8-9 paragraph 1 has been modified for clarity.}
%\commentPaul{Add verbatim the paragraph so the reviewers don't have to go searching for it.}

\item   What is the method of curing to make etched PET nGO-ODA?

\blue{The etched PET nGO-ODA was cured by oven curing.  %Oven curing has been added to the text. 
The details of oven curing are provided in the experimental section under sample preparation.}  %\commentPaul{Are details provided on the conditions for oven curing?  Please include here.}

\item     In fig 2 b the scales are not visible

\blue{Scale bars have been added.}

\item     Please elaborate how the sem images proves/confirms the etching is done because the schematic of EPET 2aii shows some structure is growing.

\blue{The schematic is meant to show additional roughness on the microfibers by etching, not growth. The SEM image Figure 2bii inset shows this additional roughness on the PET microfiber. The text on page 8-9, paragraph 1 has been modified for clarity. }
%\commentPaul{Provide the new text so the reviewers don't have to go searching for it.}

\item    Line 33 page 9 “The terephthalate group is confirmed at wavenumber cm-1.” Is some value missing here?

\blue{The value 1240 cm-1 has been added. }

\item     The FTIR  of Fig. 2c are on different scales so the stretching and other details could not be confirmed. In addition what does the lowering/enhancing of surface energy signifies with methyl groups or any other group other than chemical resistance?

\blue{The text has been modified on page 17, paragraph 1, ``Additional methyl groups from the PDMS treatment are observed at wavenumber 2850 cm$^{-1}$. Methyl groups lower the surface energy of the fabric which improves the liquid repellency property."}

\item    Fig 3a looks more of a table rather than an image. It could be marked as table instead

\blue{The authors intend for the table to be a part of Figure 3. }

\item     The different angles named in the fig 3a is not discussed much. It could be explained for a better understanding of the importance of the angles for the study of fabric behaviour.

\blue{ The following text has been added to page 10, paragraph 1, ``The static contact angle is the angle at which the contact area between the liquid and solid is unmoved. The advancing angle describes the angle when a droplet is wetting the surface and the receding angle is the angle when a droplet is dewetting the surface. The hysteresis describes how easily a droplet rolls off the surface and is calculated by the difference between the advancing and receding contact angles."}

\item    what kind of transition is Cassie-Baxter to Wenzel wetting state and how effective it is and the cause behind this transition and its role in demonstrating anti virofouling properties?

\blue{The following text has been added on page 11, paragraph 1, 
``Untreated PET and etched PET samples are fully wetting with water droplets; faster liquid spreading was observed for etched PET fibers.
After treatment, the 
EPET-nGO samples are superhydrophobic, with an average water static contact angle of $156 \pm 1\degree$ and water hysteresis of $9 \pm 1\degree$.
The combination of roughened microfibers and nGO-ODA nanomaterial creates a low surface energy surface with dual micro- and nano-scale roughness, important for rendering a robust, superhydrophobic wetting property.$^{30,31}$
The Wenzel wetting state is when the liquid is in fully wetting and in full contact with the roughness of a surface. 
After etching and nGO-ODA coating, a low energy surface with both micro- and nano-scale roughness is created that promotes Cassie Baxter wetting instead of Wenzel wetting.  
The Cassie-Baxter wetting state is when liquid sits on top of the roughness of a surface, resulting in air pockets between the surface and the liquid. 
The low contact fraction area between the surface and liquid results in high static contact angle as well as low contact angle hysteresis characteristic of superhydrophobic surfaces, where droplets ball up on the surface and easily roll off.  
The amount of Laplace pressure required for liquid to infiltrate the air pockets and transition from Cassie-Baxter  to Wenzel wetting state is called the breakthrough pressure. The breakthrough pressure is a measure of the stability of the Cassie-Baxter wetting state."}

%``The Wenzel wetting state is when the liquid is in full contact with the roughness of a surface. The Cassie-Baxter wetting state is when liquid sits on top of the roughness of a surface, resulting in air pockets between the surface and the liquid. 
%The amount of Laplace pressure required for liquid to infiltrate the air pockets and transition from Cassie-Baxter  to Wenzel wetting state is called the breakthrough pressure. The breakthrough pressure is a measure of the stability of the Cassie-Baxter wetting state."}

%``The Wenzel wetting state is when liquid is in full contact with the roughness of a surface. The Cassie-Baxter wetting state is when liquid sits on top of the roughness of a surface, resulting in air pockets between the surface and the liquid. 
%The amount of Laplace pressure required for liquid to infiltrate the air pockets and transition from Cassie-Baxter  to Wenzel wetting state is called the breakthrough pressure. Breakthrough pressure is a measure of the stability of the Cassie-Baxter wetting state and we believe anti-virofouling properties occur when a surface maintains a robust Cassie-Baxter wetting state." }

\item   Fig 3 c) Optical images of water (blue) and human saliva (yellow) droplets on i) i) untreated PET

\blue{The unnecessary i) has been removed. }

\item   What is the difference and significance of Taber rotary abrasion cycles and bleach washing durability?

\blue{The statement ``Taber rotary abrasion is a standard method for analyzing the durability of fabric coatings under mechanical wear and bleach washing is a standard protocol for re-usable, personal protective equipment in healthcare settings." has been added to Page 11 paragraph 3.}

\item   Page 11 line 18 “Fig .3d ………… and ultrasonic bleach washing …………… Fig .3e …….a function of Taber rotary abrasion cycles.” Revisit line 34 for the same

\blue{These typos have been corrected. }

\item  Page 11 line 39 “to dip coating …………….improves the stability of nGO-ODA on” what kind of stability is being discussed, mechanical thermal chemical?

\blue{``Wetting stability" has been added to the text. }

\item  Page 11 line 41 “Fig. 4b shows representative optical images of water” please verify.

\blue{This typo was replaced with Fig. 3e. }

\item   What is PBS and please check the namings of figure 4

\blue{``phosphate buffered saline (PBS)" has been added. }

\item   It is not clearly discussed about the EPGO or EPGOP treatment and what does the decrease in $p <0:05$ to $p <0:001$, signifies ?

\blue{The following was added to Page 13, paragraph 1,
``Mann-Whitney tests were used to assess the probability that that the fabric treatment reduces the amount of virus particles on the fabric compared to controls. 
%offer a degree of statistical certainty that the fabric treatment reduces the amount of virus particles on the fabric compared to controls. 
%Mann-Whitney tests offer a degree of statistical certainty that the fabric treatment reduces the amount of virus particles on the fabric compared to controls. 
Asterisks indicate the 
%probability with 
level of statistical significance with 
which the treatment
group is 
%the same as the 
different 
from the control group %via Mann-Whitney statistic tests 
with one asterisk for $p \leq 0.05$, two for $p < 0.01$ and three for $p < 0.001$.} %One asterisk corresponds to the level of certainty $p \leq 0.05$, two asterisks for $p < 0.01$ and three asterisks for $p < 0.001$. 
%Asterisks indicate the level of statistical significance with which the treatment
%group different from the control group via Mann-Whitney statistic tests with one asterisk for $p \leq 0.05$, two for $p < 0.01$ and three for $p < 0.001$.} %One asterisk corresponds to the level of certainty $p \leq 0.05$, two asterisks for $p < 0.01$ and three asterisks for $p < 0.001$. }

%``Mann-Whitney tests offer a degree of statistical certainty that the fabric treatment reduces the amount of virus particles on the fabric compared to controls. Asterisks indicate the statistical certainty that the control group is the same as the treatment from Mann-Whitney statistic tests, one asterisk for $p \leq 0.05$, two for $p < 0.01$ and three for $p < 0.001$." The EPGO or EPGOP was a typo that has been corrected. The decrease in p value is just a slight difference in statistical certainty.  }
%\commentPaul{statistical significance, right?}

\item   How 10\% bleach solution was chosen for bleach washing durability tests?

\blue{The following was added to the experimental section under Durability testing ``A concentration of 10\% bleach washing solution was chosen based on discussions with healthcare cleaning provider Cintas."}

\item   What is the requirement for PDMS to dilute in hexane?

\blue{The following was added to the experimental section for sample fabrication, ``Hexane was chosen as a solvent due to its low surface tension to easily wet the entire fabric surface during dip coating."}

\item       Is there any trial-and-error experiment were conducted for this study?

\blue{ This study focuses on the bleach durability of the fabric coating. There are trial and error experiments in this regard. The fabric with just nGO-ODA did not show bleach durable liquid repellent properties. The fabric was then etched prior to improve the liquid repellency durability of the coating; however, even with the additional etching step, significant color fastness was observed, shown in Figure 3e. Finally, a PDMS layer was added to improve the bleach wash durability by adding chemical resistance. }

\item       Errors in the figure name, and are not arranged properly according to the content in result and discussion part, it makes difficult to read.
with respect to electrical, thermal, or mechanical stability?

\blue{Figure names have been changed accordingly.}

\item  Graph scale is not uniform, difficult to identify the variations in the bar graphs.

\blue{The graph scale is uniform for each type of virus. The titers were different for each virus as each type of virus has a different level of viability and infectivity. }

\end{enumerate}

\newpage
\textbf{Reviewer: 3}

\textbf{Recommendation:} Major revisions needed as noted.

In the manuscript entitled “Coal-Derived Functionalized Graphene Oxide for Bleach Washable, Durable Anti-Virofouling Fabric Coatings”, the author reported the personal protective equipment material with the antiviral resistant. The author claiming that octadecylamine functionalized graphene-oxide/PDMS based PET fabric shows the good repellency and durability with water, human saliva and viruses. However certain aspects the authors have to modify the manuscript to improve the scientific quality.

\blue{We thank the reviewer for his or her constructive comments.
We have made improvements based on the reviewer's comments as shown below.}

\begin{enumerate}
\item The author didn’t explain properly the role of each material in the composite fabrics. The author have to elaborately explain the role of the materials like octadecylamine functionalization with GO, PET-nGO-ODA fabric and PET nGO-ODA/PDMS composite fabric.

### 2022-01-20 Latex/Main.tex
Updated document

\author{Anthony J. Galante}
\affiliation[University of Pittsburgh]
{Department of Industrial Engineering, University of Pittsburgh, Pittsburgh, PA 15261, USA}
%\altaffiliation{A shared footnote}
%\author{Fred T. Secondauthor}
%\altaffiliation{Current address: Some other place, Othert\"own,
%Germany}

@@

\author{Brady Pilsbury}
\affiliation[University of Pittsburgh]
{Department of Industrial Engineering, University of Pittsburgh, Pittsburgh, PA 15261, USA}

\author{Melbs LeMieux}
\affiliation{7901 East Riverside Drive, Bldg 1,Unit 150, Austin TX  78744}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title
  {Bleach-durable reactive silver ink coatings for anti-viral, 
  %virucidal, 
  repellent, and reusable textiles}
% \title
%   {Comparison of virus killing durability after bleach washing of superhydrophobic, silver ink and silver nanoparticle coatings on PET for improving reusable PPE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{SEM,FTIR, PPE, GQDs, GO,  PET, NETL, ODA, PDMS}
\keywords{silver,killing, deactivation,anti-virofouling, virus, self-cleaning, superhydrophobic, reusable, personal protective equipment (PPE), fabric \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \setlength {\marginparwidth }{2cm}
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

@@

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%During the most recent virus pandemic, the shortage of personal protective equipment (PPE) became a global health crisis. 
There is a major need to improve reusable personal protective equipment (PPE) with %virus inhibiting 
anti-viral
and liquid repelling functionality. %reusability  and durability for economical and environmental concerns. 
Silver is a popular coating material to incorporate virus inactivation functionality to PPE. This work consists of utilizing reactive silver inks and low surface energy polymers to achieve 
anti-viral and 
%enveloped virus inhibiting 
and liquid repellency functionalities even after ultrasonic bleach washing. The virus inhibition and repellency functionalities of reactive silver inks and silver nanoparticles on fabric are measured before and after bleach washing. The potential of reusability for reactive silver inks is compared with traditional silver nanoparticle coatings on fabric.

@@

@@

Personal protective equipment (PPE) such as  gowns,  masks, and scrubs are essential for protecting healthcare professionals from contact with bacteria and viruses that lead to infection.
% Viruses may also be transmitted via 
% respiratory droplets which can land on
%Respiratory droplets may also 
However, PPE may become contaminated during use and inadvertently contribute to the transmission of microbes \cite{Sakaguchi:10}. There is a need for textile coatings that provide PPE with better protection from infection 
%barrier properties by 
by not only repelling fluids such as 
respiratory droplets, but 
killing or deactivating microbes. In particular, there is great interest in textile coatings that provide these functionalities in reusable PPE even after laundering \cite{Kraay:18}.

Single-use or disposable PPE place a heavy burden on manufacturing supply chains that may lead to shortages during pandemics or times of great demand.\cite{cohen_contributing_2020}  
%\commentPaul{Find reference}
Single use or disposable PPE also lead to large amounts of waste and environmental impact \cite{Overcash:12}.
%Reusable PPE also 
%such as  gowns,  masks,  scrubs,  and  sheets  
%address these issues, 
Reusable PPE help alleviate burdens on manufacturing supply chains and decrease the amount of waste.  Furthermore, they are more economical on a cost per use basis.  \cite{Overcash:12} %\commentPaul{Find citation for this.}
However, one of the main challenges with reusable PPE is ensuring that they continue to provide protection from microbes after repeated use and decontamination.  Reusable PPE are decontaminated after each use by laundering with bleach.  
%are  subject to laundering with bleach for decontamination.  
% about 10\%  bleach  washing  for  20  min before being returned.\cite{cintas}
% There is a need for 
% improvements on the functionality and reusability of PPE for public health safety.\cite{Kraay:18} %Self-cleaning, medical coatings are a proactive strategy to prevent surface contamination from human body fluid, bacteria, and viruses.
% Reusable  PPE in the healthcare sector such as  gowns,  masks,  scrubs,  and  sheets  are  subject to  about 10\%  bleach  washing  for  20  min before being returned.\cite{cintas}
%, according  to  Cintas,  a  major  healthcare laundering  company. 
Therefore, virus inactivating surface coatings for reusable PPE in medical settings must be able to withstand multiple laundering cycles in bleach solution.

@@

%Silver ink is a promising alternative for fabric treatments; however, the functionality durability of silver treated fabrics after bleach washing has not been investigated. 
% The use of silver for functional PPE applications must be durable so that 
% functionality is retained after repeated bleach washing.  
This work studies the virucidal activity of
PET textiles coated with a silver amine complex ink \cite{Walker:12}
on herpes simplex viruses (HSV-1).  
PET fabric material is commonly used for medical and healthcare applications such as gowns, scrubs and caps. \cite{Rigby:97} 
HSV-1 is an enveloped virus about 160 nm in diameter that can cause cold sores and blinding herpetic eye infections.\cite{wald:07} 
The attachment and entry of HSV-1 into cells requires the interaction between the viral envelope glycoproteins and cell surface heparan sulfate (HS).\cite{baram-pinto_2009} Silver mimics HS and competes for binding sites of the virus to the cell, inhibiting the virus from cell attachment and entry.\cite{baram-pinto_2009,Galdiero:11}

>>>>

\author{Anthony J. Galante}
\affiliation[University of Pittsburgh]
{Department of Industrial Engineering, University of Pittsburgh, Pittsburgh, PA 15261, USA}
%\altaffiliation{A shared footnote}
%\author{Fred T. Secondauthor}
%\altaffiliation{Current address: Some other place, Othert\"own,
%Germany}
\author{Brady C. Pilsbury}
\affiliation[University of Pittsburgh]
{Department of Industrial Engineering, University of Pittsburgh, Pittsburgh, PA 15261, USA}

@@

\author{Melbs LeMieux}
\affiliation[Electroninks Inc]{Electroninks Inc, 7901 East Riverside Drive, Bldg 1,Unit 150, Austin TX  78744}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title
  {Reactive silver inks for antiviral, 
  %virucidal, 
  repellent medical textiles with ultrasonic bleach washing durability compared to silver nanoparticles}
% \title
%   {Comparison of virus killing durability after bleach washing of superhydrophobic, silver ink and silver nanoparticle coatings on PET for improving reusable PPE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{SEM,FTIR, XRD, PPE, PET, PDMS, PFU, PBS, HSV}
%\keywords{silver,killing, deactivation,anti-virofouling, virus, self-cleaning, superhydrophobic, reusable, personal protective equipment (PPE), fabric \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \setlength {\marginparwidth }{2cm}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\begin{tocentry}

@@

\end{tocentry}
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%During the most recent virus pandemic, the shortage of personal protective equipment (PPE) became a global health crisis. 
% There is a major need to improve reusable personal protective equipment (PPE) with %virus inhibiting 
% anti-viral
% and liquid repelling functionality. %reusability  and durability for economical and environmental concerns. 
% Silver is a popular coating material to incorporate virus inactivation functionality to PPE. 
Medical textiles are subject to particularly harsh disinfection  procedures 
%due to the 
%need to create sterile surfaces 
in healthcare settings where exposure risks are high.  
%high risks of exposure to microbes in healthcare settings.  
This work demonstrates a fabric treatment consisting of 
a reactive silver ink and low surface energy polymer that 
provides for 
superhydrophobicity and antiviral properties against enveloped virus stocks 
even after extended ultrasonic bleach washing.
%as well as 
%bleach and ultrasonic cleaning.
%wash 
%durability.  
%a bleach wash durable fabric treatment with . 
The fabric treatment 
exhibits high static contact angles and low contact angle hysteresis with water, even after 300 minutes of ultrasonic bleach washing. Similarly, after this bleach washing treatment, the fabric treatment shows reductions of infectious virus quantities by about 2 logs compared to controls for enveloped viruses The use of silver ink provides for 
better antiviral efficacy and durability compared to silver nanoparticles due to the use of reactive ionic silver, which demonstrates more conformal coverage of fabric microfibers and better %microfiber 
adhesion.  This study provides insights for improving the wash durability of antiviral silver fabric treatments and demonstrates a bleach wash durable, repellent antiviral treatment for reusable, functional personal protective equipment applications.

% This work consists of utilizing reactive silver inks and low surface energy polymers to achieve 
% anti-viral and 
% %enveloped virus inhibiting 
% and liquid repellency functionalities even after ultrasonic bleach washing. The virus inhibition and repellency functionalities of reactive silver inks and silver nanoparticles on fabric are measured before and after bleach washing. The potential of reusability for reactive silver inks is compared with traditional silver nanoparticle coatings on fabric.

@@

KEYWORDS:  virus, superhydrophobic, repellency, fabric, silver, bleach, durable, antiviral

@@

Personal protective equipment (PPE) such as  gowns,  masks, and scrubs are essential for protecting healthcare professionals from contact with bacteria and viruses that lead to infection.
% Viruses may also be transmitted via 
% respiratory droplets which can land on
%Respiratory droplets may also 
% However, PPE may become contaminated during use and inadvertently contribute to the transmission of microbes \cite{Sakaguchi:10}. 
There is a need for textile coatings that provide PPE with better protection from infection 
%barrier properties by 
by not only repelling fluids such as 
respiratory droplets, but 
also killing or deactivating microbes. In particular, there is great interest in fabric coatings that provide these protective functionalities in PPE even after repeated use and laundering \cite{Kraay:18,galante:2020,galante:2022}.
% Single-use or disposable PPE place a heavy burden on manufacturing supply chains that may lead to shortages during pandemics or times of great demand.\cite{cohen_contributing_2020}  
% %\commentPaul{Find reference}
% Single-use or disposable PPE also lead to large amounts of waste and environmental impact \cite{Overcash:12}.
% %Reusable PPE also 
% %such as  gowns,  masks,  scrubs,  and  sheets  
% %address these issues, 
% Reusable PPE help alleviate burdens on manufacturing supply chains and decrease the amount of waste.  Furthermore, they are more economical on a cost per use basis.  \cite{Overcash:12} %\commentPaul{Find citation for this.}
%One of the main challenges with reusable PPE is ensuring that they continue to provide protection from microbes after repeated use. %and decontamination.  
Furthermore, there is need for new fabric coatings that can withstand particularly harsh disinfection procedures such as high concentrations of bleach and ultrasonic cleaning that may be used in high%er 
risk settings such as healthcare.

% must withstand multiple laundering cycles in bleach solution, which is commonly used for decontamination in medical settings.

% Therefore, virus inactivating surface coatings for reusable PPE in medical settings must be able to withstand multiple laundering cycles in bleach solution. 
%Reusable PPE are decontaminated after each use by laundering with bleach.  
%are  subject to laundering with bleach for decontamination.  
% about 10\%  bleach  washing  for  20  min before being returned.\cite{cintas}
% There is a need for 
% improvements on the functionality and reusability of PPE for public health safety.\cite{Kraay:18} %Self-cleaning, medical coatings are a proactive strategy to prevent surface contamination from human body fluid, bacteria, and viruses.
% Reusable  PPE in the healthcare sector such as  gowns,  masks,  scrubs,  and  sheets  are  subject to  about 10\%  bleach  washing  for  20  min before being returned.\cite{cintas}
%, according  to  Cintas,  a  major  healthcare laundering  company.

@@

%Silver ink is a promising alternative for fabric treatments; however, the functionality durability of silver treated fabrics after bleach washing has not been investigated. 
% The use of silver for functional PPE applications must be durable so that 
% functionality is retained after repeated bleach washing.  
This work studies the virucidal activity of
polyethylene terephthalate (PET) knit fabrics coated with a silver amine complex ink \cite{Walker:12}
against herpes simplex viruses (HSV-1). 
%Once the silver ink is cured on the fabric, the additional ingredients \commentPaul{Be more specific} \commentsAnthony{These ingredients are proprietary. We could take this sentence out.} of the ink evaporate to leave a pure silver film.   
%The ink is cured thermally
PET fabric material is commonly used for medical and healthcare applications such as gowns, scrubs and caps. \cite{Rigby:97} 
HSV-1 is an enveloped virus about 160 nm in diameter that can cause cold sores and blinding herpetic eye infections.\cite{wald:07} 
The attachment and entry of HSV-1 into cells require the interaction between the viral envelope glycoproteins and cell surface heparan sulfate (HS).\cite{baram-pinto_2009} Silver mimics HS and competes for binding sites of the virus to the cell, inhibiting the virus from cell attachment and entry.\cite{baram-pinto_2009,Galdiero:11}
